---
title: "Get started with rewind"
output: rmarkdown::html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE # Shiny apps can't run inside a static document.
)
```

Shiny has bookmarking. It captures a state that you can send as a link.
Shiny also has `reactlog`. It lets *you* replay a session while you debug.
But Shiny has no undo. Users expect undo, because each other application
has it. `rewind` adds undo in one line.

## Install

Install `rewind` from CRAN:

```{r}
install.packages("rewind")
```

To get the development version from GitHub:

```{r}
# install.packages("remotes")
remotes::install_github("tenmeh/rewind")
```

## The one-line version

```{r}
library(shiny)
library(rewind)

ui <- fluidPage(
  rewind_buttons(),
  selectInput("region", "Region", c("North", "South", "East", "West")),
  sliderInput("year", "Year", 2018, 2026, 2024),
  plotOutput("plot")
)

server <- function(input, output, session) {
  rewind_enable() # <- that's it

  output$plot <- renderPlot(plot_for(input$region, input$year))
}

shinyApp(ui, server)
```

`Ctrl+Z` now walks backwards through the user's filter choices. `Ctrl+Shift+Z`
walks forwards. Every input in the session is captured automatically, with a
short list of exclusions (action buttons, `fileInput()`, and rewind's own
inputs) covered in `?rewind_enable`.

## One movement of a slider is one undo step, not forty

`sliderInput` sends an input change for each pixel of a movement. Without
grouping, you would press `Ctrl+Z` forty times to return to the start.
`rewind` puts the changes that occur within `coalesce_ms` of each other into
one history entry. The default is 400 ms.

```{r}
rewind_enable(coalesce_ms = 600) # make larger groups
```

Increase `coalesce_ms` for widgets that send changes very quickly. Decrease
it for smaller undo steps.

## Grouping several changes into one step

A "reset filters" button changes four inputs together. It must make one
undo entry, with a label that a person wrote. It must not make four
entries, and it must not use an automatic label such as "4 values changed".
Put the block inside `rewind_step()`:

```{r}
observeEvent(input$reset, {
  rewind_step(label = "Reset filters", {
    updateSelectInput(session, "region", selected = "All")
    updateSliderInput(session, "year", value = c(2018, 2026))
    updateCheckboxInput(session, "active_only", value = FALSE)
  })
})
```

## Undo for state that is not an input

`rewind` cannot see the values that you keep in `reactiveValues()`. These
can be a list of pinned items, a current selection, or a cache that you want
to undo. Register them with `rewind_track()`:

```{r}
server <- function(input, output, session) {
  rewind_enable()

  state <- reactiveValues(pinned = character(0))
  rewind_track(state, fields = "pinned")

  observeEvent(input$pin, {
    state$pinned <- c(state$pinned, input$item)
  })
}
```

An undo now restores the pin list. It also restores the inputs that changed
at the same time.

## A visual history rail

`rewind_ui()` draws the stack as a scrubbable list. Click a step to move to
it. The labels come from the values that changed. The rail thus shows
"region, year" and not "state 7":

```{r}
ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(rewind_ui()),
    mainPanel(plotOutput("plot"))
  )
)
```

## Modules

You can call `rewind_enable()` inside a `moduleServer()`. `rewind` captures
the inputs with their module-local names. It adds the namespace with
`session$ns()` at a restore. This is the same as each other operation on
Shiny inputs inside a module.

All modules share `session$userData`. Call `rewind_enable()` once, at the
position in the module tree that is best for your application. A second
call in a different module uses the same history. It does not make a second
history.

## Stop capture around changes that your code makes

Your code can restore a saved session, or apply a URL bookmark at the start.
These changes must not become an undo step. Put that code between
`rewind_pause()` and `rewind_resume()`:

```{r}
server <- function(input, output, session) {
  rewind_enable()

  rewind_pause()
  # ... apply a saved bookmark here ...
  rewind_resume()
}
```

Use `rewind_disable()` to stop undo and redo completely during a session.
An example is a user role that the application knows only after the session
starts.

## What to read next

- `?rewind_enable` for the full list of the inputs that `rewind` captures
- The [README](https://github.com/tenmeh/rewind#readme) for the mechanics of
  how restore works and the echo problem it has to avoid
- The [reference index](../reference/index.html) for every exported function
