---
title: "Get started with rewind"
output: rmarkdown::html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE # Shiny apps can't run inside a static document.
)
```

Shiny has bookmarking, which captures a state you can link to. It has
`reactlog`, which lets *you* replay a session while debugging. It has no
undo - the thing every user already expects from every other application
they use. `rewind` adds it in one line.

## Install

`rewind` isn't on CRAN yet. Install the development version from GitHub:

```{r}
# install.packages("remotes")
remotes::install_github("tenmeh/rewind")
```

## The one-line version

```{r}
library(shiny)
library(rewind)

ui <- fluidPage(
  rewind_buttons(),
  selectInput("region", "Region", c("North", "South", "East", "West")),
  sliderInput("year", "Year", 2018, 2026, 2024),
  plotOutput("plot")
)

server <- function(input, output, session) {
  rewind_enable() # <- that's it

  output$plot <- renderPlot(plot_for(input$region, input$year))
}

shinyApp(ui, server)
```

`Ctrl+Z` now walks backwards through the user's filter choices. `Ctrl+Shift+Z`
walks forwards. Every input in the session is captured automatically, with a
short list of exclusions (action buttons, `fileInput()`, and rewind's own
inputs) covered in `?rewind_enable`.

## Dragging a slider is one undo step, not forty

`sliderInput` fires an input change on every pixel of a drag. Without
coalescing, undoing that drag would take forty presses of `Ctrl+Z` to get
back to where you started. `rewind` groups changes that land within
`coalesce_ms` (400ms by default) of one another into a single history entry:

```{r}
rewind_enable(coalesce_ms = 600) # group more aggressively
```

Raise `coalesce_ms` for widgets that fire very rapidly, or lower it if you
want finer-grained undo steps.

## Grouping several changes into one step

A "reset filters" button that changes four inputs at once deserves one undo
entry, with a label a human wrote - not four entries, and not an
auto-generated label like "4 values changed". Wrap the block in
`rewind_step()`:

```{r}
observeEvent(input$reset, {
  rewind_step(label = "Reset filters", {
    updateSelectInput(session, "region", selected = "All")
    updateSliderInput(session, "year", value = c(2018, 2026))
    updateCheckboxInput(session, "active_only", value = FALSE)
  })
})
```

## Undoing state that isn't an input

Anything you keep in `reactiveValues()` - a pinned-items list, a running
selection, a computed cache you want to be undoable - is invisible to
`rewind` until you register it with `rewind_track()`:

```{r}
server <- function(input, output, session) {
  rewind_enable()

  state <- reactiveValues(pinned = character(0))
  rewind_track(state, fields = "pinned")

  observeEvent(input$pin, {
    state$pinned <- c(state$pinned, input$item)
  })
}
```

Undo now rolls the pin list back along with whatever inputs changed at the
same time.

## A visual history rail

`rewind_ui()` renders the stack as a scrubable list. Click any step to jump
straight to it. Labels come from what actually changed, so the rail reads
"region, year" rather than "state 7":

```{r}
ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(rewind_ui()),
    mainPanel(plotOutput("plot"))
  )
)
```

## Modules

`rewind_enable()` works inside a `moduleServer()`. Inputs are captured under
their module-local names and re-qualified with `session$ns()` on restore, the
same as any other Shiny input handling inside a module. `session$userData` is
shared across modules, so call `rewind_enable()` once, wherever in the module
tree suits the app - calling it again elsewhere reuses the same session-wide
history rather than creating a second, independent one.

## Suspending capture around programmatic changes

Restoring a saved session or applying a URL bookmark on startup shouldn't
itself become an undo step the user can step back into. Wrap that code in
`rewind_pause()` / `rewind_resume()`:

```{r}
server <- function(input, output, session) {
  rewind_enable()

  rewind_pause()
  # ... apply a saved bookmark here ...
  rewind_resume()
}
```

For turning undo/redo off entirely partway through a session - for example,
gated behind a user role decided after the session starts - see
`rewind_disable()`.

## What's next

- `?rewind_enable` for the full list of what is and isn't captured by default
- The [README](https://github.com/tenmeh/rewind#readme) for the mechanics of
  how restore works and the echo problem it has to avoid
- The [reference index](../reference/index.html) for every exported function
